Community engagement and subgroup meta-knowledge: Some factors in the soul of a community

### 2015-08-05 PaperFinalDraft.Rnw
dplyr-ized citystateResp to close #8

<<YN2010>>=
dfall <- data.frame()
for (i in 1:length(twolevel)){
  tmp <- data.frame(Question=twolevel[i], prop.table(table(sotc10[,twolevel[i]])))
  dfall <- rbind(dfall, tmp)
}
names(dfall) <- c("Question", "Response", "Freq")

citystateResp <- function(dataset){
  # Create a loop that calculates the percentages of cases in each factor level for each city
  dfcities <- data.frame()
  questionsIn <- which(twolevel %in% names(dataset))
  for (j in 1:length(levels(dataset$citystate))){
    thiscity <- subset(dataset, citystate==levels(dataset$citystate)[j])
    for (i in questionsIn){
      tmp <- data.frame(levels(dataset$citystate)[j], Question=twolevel[i], prop.table(table(thiscity[,twolevel[i]])))
      dfcities <- rbind(dfcities, tmp)
    }
  }
  names(dfcities) <- c("citystate", "Question", "Response", "Freq")

  # Now we calculate difference between the overall and the cities to see where the variation comes from
  dfcities$diff <- 0
  for (i in 1:length(levels(dfcities$Question))){
    # Yes
    overallrate <- dfall[dfall$Question==levels(dfall$Question)[i] & dfall$Response=="Yes",]$Freq
    dfcities$diff[(dfcities$Question==levels(dfcities$Question)[i]& dfcities$Response=="Yes")] <- (overallrate - dfcities[(dfcities$Question==levels(dfcities$Question)[i] & dfcities$Response=="Yes"),]$Freq)/overallrate
  }
  return(dfcities)
}

dfcities10 <- citystateResp(sotc10)
dfcitiesYes <- dfcities10[dfcities10$Response=="Yes",]
dfcitiesYes$col <- dfcitiesYes$dif>0
dfcitiesYes$Question <- factor(dfcitiesYes$Question, levels=levels(dfcitiesYes$Question)[order(dfall[dfall$Response=="Yes",]$Freq)])
@

<<startingoverall>>=
dfcities09 <- citystateResp(sotc09) %>% 
  mutate(year="2009")
dfcities08 <- citystateResp(sotc08) %>% 
  mutate(year="2008")
dfcities10 <- dfcities10 %>% 
  mutate(year="2010")
md1 <-rbind(dfcities10, dfcities09, dfcities08) %>% 
  filter(Response=="Yes") %>% 
  select(-Response, -diff) %>% 
  filter(Question %in% c("registered_vote", "voted_local", "volunteer", "worked_change", "public_meeting"))
md1 <- melt(md1)
md1 <- dcast(md1, citystate + year ~ Question, mean)
md1$year <- as.numeric(md1$year)
masterData <- left_join(masterData, md1)
@

>>>>

<<YN2010>>=
dfcities10 <- sotc10 %>%
  select(citystate, one_of(twolevel)) %>%
  mutate(id = 1:n()) %>%
  group_by(citystate) %>%
  tidyr::gather("Question", "Response", -id, -citystate) %>%
  group_by(citystate, Question, Response) %>%
  filter(!is.na(Response)) %>%
  summarise(n = n()) %>%
  mutate(Freq = n/sum(n)) %>%
  filter(Response=="Yes") %>%
  merge(select(filter(df10, Response=="Yes"), Freq), by="Question") %>%
  mutate(diff=Freq.y-Freq.x) %>%
  select(-Freq.y, -n, Freq=Freq.x) %>%
  arrange(citystate) %>%
  mutate(year="2010")

dfcitiesYes <- dfcities10 %>%
  filter(Response=="Yes") %>%
  mutate(col = diff>0)

dfcitiesYes$Question <- factor(dfcitiesYes$Question, levels=levels(dfcitiesYes$Question)[order(dfall[dfall$Response=="Yes",]$Freq)])
@

<<startingoverall>>=
dfcities09 <- sotc09 %>%
  select(citystate, one_of(twolevel[1:5])) %>%
  mutate(id = 1:n()) %>%
  group_by(citystate) %>%
  tidyr::gather("Question", "Response", -id, -citystate) %>%
  group_by(citystate, Question, Response) %>%
  filter(!is.na(Response)) %>%
  summarise(n = n()) %>%
  mutate(Freq = n/sum(n)) %>%
  filter(Response=="Yes") %>%
  merge(select(filter(df10, Response=="Yes"), Freq), by="Question") %>%
  mutate(diff=Freq.y-Freq.x) %>%
  select(-Freq.y, -n, Freq=Freq.x) %>%
  arrange(citystate) %>%
  mutate(year="2009")

dfcities08 <- sotc08 %>%
  select(citystate, one_of(twolevel[1:5])) %>%
  mutate(id = 1:n()) %>%
  group_by(citystate) %>%
  tidyr::gather("Question", "Response", -id, -citystate) %>%
  group_by(citystate, Question, Response) %>%
  filter(!is.na(Response)) %>%
  summarise(n = n()) %>%
  mutate(Freq = n/sum(n)) %>%
  filter(Response=="Yes") %>%
  merge(select(filter(df10, Response=="Yes"), Freq), by="Question") %>%
  mutate(diff=Freq.y-Freq.x) %>%
  select(-Freq.y, -n, Freq=Freq.x) %>%
  arrange(citystate) %>%
  mutate(year="2008")

md1 <-rbind(dfcities10, dfcities09, dfcities08) %>% 
  select(-Response, -diff) %>% 
  filter(Question %in% c("registered_vote", "voted_local", "volunteer", "worked_change", "public_meeting"))
md1 <- melt(md1)
md1 <- dcast(md1, citystate + year ~ Question, mean)
md1$year <- as.numeric(md1$year)
masterData <- left_join(masterData, md1)
@
